---
title: "Co-Authorship Network Analysis - sample"
author: "Amir Imani"
date: "1 August 2016"
output: pdf_document
---

Having +6000 articles sitting on my hard drive, I thought to play around with them. 
The idea is that to see how tightly knitted the network of scholars who write about **Syntethic Biology** and **Ethics**. 

To do so, I picked up a random sample of 85 articles with more than one author to creat my network. Each node in the network is an author, and directed edges go from all the authors of one articles to each others. 

relying on R, I have extracted author names from articles, created a matrix of co-authorship, transformed the matrix to an edge list, and conducted network analysis.


```{r, echo=FALSE}
library(reshape)
library(dplyr)
library(stringr)
library(stringdist)

Matrix2Edge <- function(x){  
  data <- x
  final <- data.frame("a"=character(),"b"=character()) 
  #the idea is to iterate over the columns of  the dataframe, melt it, and each time drop the melted column
  for (i in 1:(ncol(source_file)-1)){
    id = paste("V", i, sep="")
    edge_list <- melt(data, id = (id))
    data[id] <- NULL
    keeps <- c(id, "value")
    edge_list <- edge_list[keeps]
    colnames(edge_list) <- colnames(final)
    final <- rbind(final,edge_list)
  }
  final <- final[!(final$b=="" | final$b==" "),]
  return(final)
}

simmilarity_index <- function(x, y){ 
  sim_index <- 1- stringdist(x, y, method = "lv")/max(nchar(as.character(x)), nchar(as.character(y)))
  return(sim_index)  
}
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
