---
output: word_document
---


Having +6000 articles sitting on my hard drive, and a few days off from work, I thought it could be interseting to see how the scholars are related to each other in the field of **Syntethic Biology**. 

I have done an analysis on **co-authorship** to find the dynamic among scholars, to assess collaboration trends and to identify leading scientists and organizations. The analysis reveals the social structure of the networks by identifying actors and their connections. 

To do so, I picked up a random sample of 85 articles with more than one author to creat my network, scraped the authors, assign each as a node in the network, and directed edges go from all the authors of one articles to each others. 


```{r, echo=FALSE, eval=FALSE}
#load the files
edge_list = read.csv("edge_list_synthbio.csv", sep = ",", header = T)
edge_list$X <- NULL

#creating the graph
library(igraph)
el=as.matrix(edge_list)
graph=graph.edgelist(el,directed=TRUE)
density = graph.density(graph)

```


The basic steps for conducting co-authorship studies in health research are described and common network metrics are presented. 

As seen from the grapg density, it is a sparse graph.

```{r, echo=FALSE, eval=TRUE}
print(density)
```
